---
title: "Investigate Database Recovery"
name: "Vincent Tam"
course: "CS5200"
semester: "Fall 2022"
output: pdf_document
---
```{r}
# Import RSQLite library
library(RSQLite)

# Load DBI library for functions (dbConnect, dbListTables, dbWriteTable)
library(DBI)

dbfile = "example.sqlite"

#create dbcon variable
dbconLite <- dbConnect(RSQLite::SQLite(), dbfile)
```

**Choose a database recovery problem or scenario (perhaps from work) and then propose a solution using the techniques described in Chapter 11 in the textbook. Briefly describe the technique, when it is appropriate to use and what recovery problem it solves.**  

A classic example of a bank transfer between two bank accounts from the same database. One account is debited and the other is credited with funds. Assume that an error occurred in between the transfer and database recovery is necessary.  

One solution is to use log-based recovery.(Date, C. J., et al. Introduction to Database Systems. Pearson Education, 2006.).The database keeps logs of all transactions and when an error is detected. The database can revert back to a previous state prior to the failed transaction. For example if Person A had $1000 and was to pay Person B $500. If the transaction failed, the database can just reflect the same state it is in now, where Person A will still have $1000 in his/her account and the payment to Person B will not have been processed.  

```{sql connection=dbconLite}
/* Enable foreign key constraints */

PRAGMA foreign_keys = ON
```

```{sql connection=dbconLite}
/* Create Author Table */
CREATE TABLE Author (
  aid NUMERIC NOT NULL,
  name TEXT NOT NULL,
  email TEXT NOT NULL,
  affiliation TEXT CHECK(affiliation IN ('PT','FT','Contract','Associate')),
  PRIMARY KEY (aid)
)
```

```{sql connection=dbconLite}
/* Create Modules Table */
CREATE TABLE Module (
  number TEXT NOT NULL,
  title TEXT NOT NULL,
  lengthInMin NUMERIC NOT NULL,
  PRIMARY KEY (number),
  FOREIGN KEY (number) REFERENCES Author(aid)
)
```

```{sql connection=dbconLite}
/* Insert into Author table. */

INSERT INTO Author (aid,name,email,affiliation) VALUES 
  (1,"Jose","Jose@northeastern.edu","PT"),
  (2,"Dan","Dan@northeastern.edu","Contract"),
  (3,"Martin","Martin@northeastern.edu","FT"),
  (4,"Kathleen","Kathleen@northeastern.edu","Associate");
```

```{sql connection=dbconLite}
/* Insert into Modules table. */

INSERT INTO Module (number,title,lengthInMin) VALUES
  ("1","Introduction",20),
  ("2","Test 1", 30),
  ("3","Test 2", 120),
  ("4","Test 3", 360);
```

```{sql connection=dbconLite}
/* Using Transactions, update the Author and Module Table concurrently to prevent errors */

BEGIN TRANSACTION;

UPDATE Module
   SET title = "New Intro"
   WHERE number = "1";
   
UPDATE Author
   SET name = "Vincent", email = "Vincent@northeastern.edu"
   WHERE aid = 1;
   
COMMIT TRANSACTION;
```

```{sql connection=dbconLite}
/* Verify that transaction was completed in Authors table. */
SELECT * FROM Author;
```

```{sql connection=dbconLite}
/* Verify that transaction was completed  in Modules table. */
SELECT * FROM Module;
```

```{r}
#Disconnect From Database
dbDisconnect(dbconLite)
```